---
title: "Jolee_Markdown"
author: "Jolee Thirtyacre, Connor Whalen"
date: "2024-07-17"
output:
  html_document:
    theme: readable
    highlight: textmate
    toc: true
    toc_float: true
    toc_depth: 3
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

### General instructions for R Markdown
* For helpful instructions on how to start using R Markdown, visit [this link] (https://fish497.github.io/website/lectures/week_07/intro_rmarkdown.html).
* To insert code chunks, press command+option+i (mac), ctrl+alt+i (windows)
* Anything inside a code chunk will run as normal code, anything outside a code chunk will appear in the document as plain text. 


## Questions we are asking: 
1. Is increased parasitism associated with decreased body condition?
* This is complicated because we are proposing an observational study and the causality that we are attempting to describe can go in both directions. Fish that have high parasitism could be skinner, but also skinner fish could be more susceptible to increased parasitism. We won’t be able to parse this out using our dataset, but it is still an interesting question worth answering. We just need to be careful to not make any absolute statements based on this (which is why we use the language ‘associated’ instead of ‘affected by’). 
* There is the element of time that we can incorporate into this as well, which is to say that we could look at body condition over time and see what trends exist. If there are any non-linear trends in body condition through time (meaning it doesn’t just go down or just go up or just stay the same), we can then incorporate the variable of parasite presence/burden to see if there is any correlation between the non-linear changes in body condition and parasite counts. This would basically look like on a graph, you have a big drop in body size followed a period of time later by an increase in parasite presence in these fish, you could say something like skinnier fish are associated with an increased parasite burden, since we would see that some event caused body size to decrease and as a result there was more parasites in the fish. But again we have to be really careful to not imply causation if all we have is correlation. Chelsea said she would be surprised if we noticed this effect, but it could be worth looking at anyway. 
2. Is the type of parasite important in determining whether parasitism is associated with decreased body condition?
* Parasites exist in different parts of the body based on what they are: monogenes will always be on the gills, adult trematodes and adult cestodes will always be in the intestines, metacercariae can be most anywhere on the body, etc. 
* To answer this question, we would need to go through the data and identify what parasites are in their adult stage versus their larval stage. This would involve a bit of research into what the different (larval or adult) stages are present for each of the ones we have found. You would then be also able to look at the literature to see where that parasite would live within the fish at that stage in its life cycle, and then we could make commentaries about how different stages/types of parasites could be associated with changes to body condition. For example, if we had a ton of adult cestodes in the intestine and found that fish to have a very bad condition score, we could talk about how adult cestodes require more nutrients based on their big size, etc. 

## Methods
We removed "one-off" species as defined by appearing less than five times across all species datasets and if the parasite also exists in its own group (e.g., We found four acanthacephalean species across four fish species, so acanths were removed because less than five were found and they exist in a seperate group from other parasites. e.g., TREM.META was kept because even though it only appeared once in the H. nuchalis dataset, it will be lumped in with the other trematode metacercaria data points.)

#### Rules for discriminating between adult and larval: 
* For any parasite we have identified to a specific taxonomic level (family, genus, etc.), we will look at that life cycle in the literature and ID what stage we have. 
* Anything labeled as META or CYST are considered larval, as metacercaria and cysts are by definition non-final adult forms of parasites.  
* Anything labeled as TREM, NEM, ACANTH, MONO, COPE, or CEST are considered adults of those groups, as we were able to ID them to larger taxonomic groups because they were not in cysts. 
* lists in the code below are the complete lists of which parasites are considered adult or larval

#### Fulton's Condition Factor Equation:
| Variable | Definition                                  |
|----------|---------------------------------------------|
| K        | Fulton's Condition Factor (<1=good, >1=bad) |
| W        | weight of host fish (in grams)              |
| SL       | standard length of fish (in millimeters)    |
$$
K = W/SL^3 *100
$$

## Code for Analyses

#### Install packages
```{r install packages, eval=FALSE, echo=FALSE}
install.packages("tidyverse")
install.packages("janitor")
install.packages("ggeffects")
install.packages("rmarkdown")
install.packages("caret")
install.packages("tibble")
install.packages("DHARMa")
```

#### Load packages
```{r load packages, echo=TRUE}
library(ggeffects)
library(tidyverse)
library(ggpubr)
library(caret)
library(tibble)
library(readr)
library(dbplyr)
library(lubridate)
library(ggplot2)
library(glmmTMB)
library(DHARMa)
```

#### Load datasets 
```{r CW datasets, echo=FALSE}
P.vigilax.data <- read.csv("data/processed/Pimephales_vigilax_processed_human_readable_UPDATED_2024.07.10.csv") %>% janitor::clean_names()
I.punctatus.data <- read.csv("data/processed/Ictalurus_punctatus_processed_human_readable.csv_UPDATED_2024.07.19.csv")  %>%
  janitor::clean_names() %>%
  mutate(day_collected = as.integer(day_collected))
N.atherinoides.data <- read.csv("data/processed/Notropis_atherinoides_processed_human_readable.csv") %>% 
  janitor::clean_names()
```

## Giving adult and larval IDs so we can look at life stage compared to parasite burden
### _Pimephales vigilax_ - add columns to identify species as an adult or larval stage and new column for total parasite count
```{r p.vig a/l stage}
Pvig.life.stage <- P.vigilax.data %>%
  mutate(adult = rowSums(select(., 
                      "acanth_bigb", 
                      "acanth_bkr",
                      "nem_cona",
                      "nem_cap",
                      "nem_dich",
                      "nem_larv",
                      "nem_trbk",
                      "nem_unk",
                      "cest_godz",
                      "cest_comp",
                      "cest_ntg",
                      "cope_cl",
                      "cope_grab",
                      "cest_pea",
                      "cest_l",
                      "cest_vit",
                      "cest_unk",
                      "mono_gyro",
                      "mono_dact",
                      "mono_lg",
                      "trem_rnda",
                      "trem_unk"))) %>%
  mutate(larval = rowSums(select(.,
                      "meta_hs",
                      "meta_unk", 
                      "myx_go",
                      "myx_geom",
                      "myxo_bc",
                      "myx_thel",
                      "myxo_sbad",
                      "myxo_unk",
                      "myxo_up",
                      "trem_met",
                      "trem_sss",
                      "trem_buc",
                      "trem_metaf",
                      "trem_metags",
                      "trem_ms",
                      "trem_nea"))) %>%
  mutate(adult = as.integer(adult)) %>% 
  mutate(larval = as.integer(larval)) %>%
  mutate(psite_count = rowSums(select(.,
                                      "adult", 
                                      "larval"))) %>%
  janitor::clean_names()

```

### _Ictalurus punctatus_ - add columns to identify species as an adult or larval stage and new column for total parasite count
```{r i. pun a/l stage}
Ipun.life.stage <- I.punctatus.data %>%
  mutate(larval = rowSums(select(., 
                      "nem_cyst",
                      "trem_meta_gg",
                      "trem_meta_unk",
                      "trem_lg",
                      "nem_cyst",
                      "myx_tail",
                      "myx_geom"))) %>%
  mutate(adult = rowSums(select(.,
                      "acan_ostr",
                      "cest_comp",
                      "cest_meg",
                      "leech_kut",
                      "nem_nmts",
                      "mono_ip",
                      "mono_unk",
                      "nem_phar",
                      "nem_pty",
                      "nem_sp",
                      "nem_taf",
                      "nem_unk",
                      "nmorph",
                      "trem_2",
                      "trem_allo",
                      "trem_ble",
                      "trem_crep",
                      "trem_f",
                      "trem_lip",
                      "trem_megict",
                      "trem_smile",
                      "trem_unk"))) %>%
  select(-nmorph, -leech_kut) %>% 
  mutate(adult = as.integer(adult)) %>% 
  mutate(larval = as.integer(larval)) %>%
  mutate(psite_count = rowSums(select(.,
                                      "adult", 
                                      "larval"))) %>%
  janitor::clean_names()
```

### _Notropis athernoides_ - add columns to identify species as an adult or larval stage and new column for total parasite count
```{r n. ath a/l stage}
Nath.life.stage <- N.atherinoides.data %>%
  mutate(larval = rowSums(select(., 
                      "nem_cyst",
                      "trem_larv",
                      "trem_meta",
                      "myx_sp", 
                      "myx_round",
                      "trem_cyst"))) %>%#
  mutate(adult = rowSums(select(.,
                      "cest_comp",
                      "cest_hyd",
                      "cest_lvs",
                      "cest_tri",
                      "cope_a",
                      "mono_all",
                      "mono_unk",
                      "nem_bran",
                      "nem_cap",
                      "nem_spky",
                      "nem_tfs",
                      "nem_twas",
                      "nem_unkn",
                      "trem_gb",
                      "trem_l"))) %>%
  mutate(adult = as.integer(adult)) %>% 
  mutate(larval = as.integer(larval)) %>%
  mutate(psite_count = rowSums(select(.,
                                      "adult", 
                                      "larval"))) %>%
  janitor::clean_names()
```

#### Modify datasets to refect only parasites that are able to occur in larval and adult stages in the host. 
- we are doing this because in order to have a truly comparable dataset of parasites and their association with host body condition, we need to analyze parasites in a uniform way that does not skew the data one direction or another. For example, if you look at trematode parasites, both metacercaria and adult trematodes can live within a fish host. Now, compare that to monogenes who only occur as adults in the fish host. Because their larval stage will not occur in these fish hosts, by including them in our analyses we will be skewing the adult subset of parasites high. The monogenes have no larval counterpart in the fish, so they will always push the adult count higher. By removing monogenes from our analyses, we will have an even ground to compare larval and adult parasites. Therefore, parasites to __NOT__ include in our analyses are: 
- monogenes
- copepods
- acanthocephalans
- myxozoans

#### Mutate the data to remove the above-listed parasite groups from the analyses. 
```{r remove groups, echo=TRUE}
####
Pvig.life.stage <- P.vigilax.data %>%
  mutate(adult = rowSums(select(., 
                      "acanth_bigb", 
                      "acanth_bkr",
                      "nem_cona",
                      "nem_cap",
                      "nem_dich",
                      "nem_larv",
                      "nem_trbk",
                      "nem_unk",
                      "cest_godz",
                      "cest_comp",
                      "cest_ntg",
                      "cope_cl",
                      "cope_grab",
                      "cest_pea",
                      "cest_l",
                      "cest_vit",
                      "cest_unk",
                      "mono_gyro",
                      "mono_dact",
                      "mono_lg",
                      "trem_rnda",
                      "trem_unk"))) %>%
  mutate(larval = rowSums(select(.,
                      "meta_hs",
                      "meta_unk", 
                      "myx_go",
                      "myx_geom",
                      "myxo_bc",
                      "myx_thel",
                      "myxo_sbad",
                      "myxo_unk",
                      "myxo_up",
                      "trem_met",
                      "trem_sss",
                      "trem_buc",
                      "trem_metaf",
                      "trem_metags",
                      "trem_ms",
                      "trem_nea"))) %>%
  mutate(adult = as.integer(adult)) %>% 
  mutate(larval = as.integer(larval)) %>%
  mutate(psite_count = rowSums(select(.,
                                      "adult", 
                                      "larval"))) %>% 
  select(-matches("^(myx_|myxo_|acanth_|mono_|cope_)")) %>% 
  janitor::clean_names()

####
Ipun.life.stage <- I.punctatus.data %>%
  mutate(larval = rowSums(select(., 
                      "nem_cyst",
                      "trem_meta_gg",
                      "trem_meta_unk",
                      "trem_lg",
                      "nem_cyst",
                      "myx_tail",
                      "myx_geom"))) %>%
  mutate(adult = rowSums(select(.,
                      "acan_ostr",
                      "cest_comp",
                      "cest_meg",
                      "leech_kut",
                      "nem_nmts",
                      "mono_ip",
                      "mono_unk",
                      "nem_phar",
                      "nem_pty",
                      "nem_sp",
                      "nem_taf",
                      "nem_unk",
                      "nmorph",
                      "trem_2",
                      "trem_allo",
                      "trem_ble",
                      "trem_crep",
                      "trem_f",
                      "trem_lip",
                      "trem_megict",
                      "trem_smile",
                      "trem_unk"))) %>%
  mutate(adult = as.integer(adult)) %>% 
  mutate(larval = as.integer(larval)) %>%
  mutate(psite_count = rowSums(select(.,
                                      "adult", 
                                      "larval"))) %>% 
  select(-matches("^(myx_|acan_|mono_|leech_|nmorph)")) %>%
  janitor::clean_names()

#### 
Nath.life.stage <- N.atherinoides.data %>%
  mutate(larval = rowSums(select(., 
                      "nem_cyst",
                      "trem_larv",
                      "trem_meta",
                      "myx_sp", 
                      "myx_round",
                      "trem_cyst"))) %>%#
  mutate(adult = rowSums(select(.,
                      "cest_comp",
                      "cest_hyd",
                      "cest_lvs",
                      "cest_tri",
                      "cope_a",
                      "mono_all",
                      "mono_unk",
                      "nem_bran",
                      "nem_cap",
                      "nem_spky",
                      "nem_tfs",
                      "nem_twas",
                      "nem_unkn",
                      "trem_gb",
                      "trem_l"))) %>%
  mutate(adult = as.integer(adult)) %>% 
  mutate(larval = as.integer(larval)) %>%
  mutate(psite_count = rowSums(select(.,
                                      "adult", 
                                      "larval")))  %>% 
  select(-matches("^(myx_|mono_|cope_)")) %>%
  janitor::clean_names()
```

### Calculating Fulton's Condition Factor
```{r FCF calculated, echo=TRUE}
Nath.life.stage <- Nath.life.stage %>%
  mutate(body.index.ath = ((Nath.life.stage$weight_mg/((Nath.life.stage$standard_length_mm)^3)))*(100)) %>%
  janitor::clean_names() %>%
  filter(sex != "MA" & sex != "")

Pvig.life.stage <- Pvig.life.stage %>%
  mutate(body.index.vig = ((Pvig.life.stage$weight_mg/((Pvig.life.stage$standard_length_mm)^3)))*(100)) %>%
  janitor::clean_names() %>%
  filter(sex != "MA" & sex != "")

Ipun.life.stage <- Ipun.life.stage %>%
  mutate(body.index.pun = ((Ipun.life.stage$weight_mg/((Ipun.life.stage$standard_length_mm)^3)))*(100)) %>%
  janitor::clean_names() %>%
  filter(sex != "MA" & sex != "")

```

#### Combine datasets and scale data. Add condition factor back into the dataframes. 
```{r}
na_ip_pv_combined <- bind_rows(Nath.life.stage, Pvig.life.stage, Ipun.life.stage, .id = "dataset")
na_ip_pv_pp <- preProcess(na_ip_pv_combined, method = c("center", "scale"))
na_ip_pv_scaled <- predict(na_ip_pv_pp, na_ip_pv_combined)
notath_scaled <- na_ip_pv_scaled %>% filter(dataset == "1") %>%
  mutate(body.index.notath = ((Nath.life.stage$weight_mg/((Nath.life.stage$standard_length_mm)^3)))*(100)) %>%
  janitor::clean_names()
pimvig_scaled <- na_ip_pv_scaled %>% filter(dataset == "2") %>%
  mutate(body.index.pimvig = ((Pvig.life.stage$weight_mg/((Pvig.life.stage$standard_length_mm)^3)))*(100)) %>%
  janitor::clean_names()
ictpun_scaled <- na_ip_pv_scaled %>% filter(dataset == "3") %>%
  mutate(body.index.ictpun = ((Ipun.life.stage$weight_mg/((Ipun.life.stage$standard_length_mm)^3)))*(100)) %>%
  janitor::clean_names()
```

#### preliminary body condition index plots using original dataframe (with above-mentioned exclusions)
```{r exploratory plots, echo=FALSE, eval=FALSE}

# plot condition factor vs parasite burden in total - may be interesting
IP.psc.fcf <- ggplot(Ipun.life.stage, 
       aes(x = psite_count, y = body_index_pun, colour = ci)) + geom_point() + xlab("Parasite Count")  + 
  ylab("FCF")  + theme_bw() + stat_smooth(method=lm)

NA.psc.fcf <- ggplot(Nath.life.stage, 
       aes(x = psite_count, y = body_index_ath, colour = ci)) + geom_point() + xlab("Parasite Count")  + 
  ylab("FCF")  + theme_bw() + stat_smooth(method=lm)

PV.psc.fcf <- ggplot(Pvig.life.stage, 
       aes(x = psite_count, y = body_index_vig, colour = ci)) + geom_point() + xlab("Parasite Count")  + 
  ylab("FCF")  + theme_bw() + stat_smooth(method=lm)

ggarrange(IP.psc.fcf, NA.psc.fcf, PV.psc.fcf, common.legend = TRUE, legend = "bottom", labels = c("A", "B", "C"))
# more variability than the others

```

### Plot life stage data using original dataframe (with above-mentioned exclusions)
```{r life stage plots, echo=TRUE}
# plot condition factor vs binary code for life stage (aka "stage-specific parasite stuff")
NA.bc.a <- ggplot(Nath.life.stage,
       aes(x = adult, y = body_index_ath, color = sex)) + xlab("Adult")  + geom_point() +
  ylab("Fulton's Condition Factor")  + theme_bw() + stat_smooth(method=lm)

NA.bc.l <- ggplot(Nath.life.stage,
       aes(x = larval, y = body_index_ath, color = sex)) + xlab("Larval")  + geom_point() +
  ylab("Fulton's Condition Factor")  + theme_bw() + stat_smooth(method=lm)

# plot condition factor vs binary code for life stage (aka "stage-specific parasite stuff")
IP.bc.a <- ggplot(Ipun.life.stage,
       aes(x = adult, y = body_index_pun, color = sex)) + xlab("Adult")  + geom_point() + 
  ylab("Fulton's Condition Factor")  + theme_bw() + stat_smooth(method=lm)

IP.bc.l <- ggplot(Ipun.life.stage, 
       aes(x = larval, y = body_index_pun, color = sex)) + xlab("Larval")  + geom_point() +
  ylab("Fulton's Condition Factor")  + theme_bw() + stat_smooth(method=lm)

# plot condition factor vs binary code for life stage (aka "stage-specific parasite stuff")
PV.bc.a <- ggplot(Pvig.life.stage, 
       aes(x = adult, y = body_index_vig, color = sex)) + xlab("Adult")  + geom_point() +
  ylab("Fulton's Condition Factor")  + theme_bw() + stat_smooth(method=lm)

PV.bc.l <- ggplot(Pvig.life.stage,
       aes(x = larval, y = body_index_vig, color = sex)) + xlab("Larval")  + geom_point() +
  ylab("Fulton's Condition Factor")  + theme_bw() + stat_smooth(method=lm)

ggarrange(IP.bc.a, IP.bc.l, NA.bc.a, NA.bc.l, PV.bc.a, PV.bc.a, common.legend = TRUE, legend = "bottom", labels = c("A", "B", "C", "D", "E", "F"))
# more variability than the others

## double check sex, weights
## should we exclude these outliers

```

#### preliminary body condition index plots using scaled dataframe (with above-mentioned exclusions)
```{r exploratory plots scaled, echo=FALSE, eval=FALSE}

# plot condition factor vs parasite burden in total - may be interesting
IP.psc.fcf <- ggplot(ictpun_scaled, 
       aes(x = body_index_ictpun, y = psite_count, colour = ci)) + geom_point() + xlab("FCF")  + 
  ylab("psite count")  + theme_bw() + stat_smooth(method=lm)

NA.psc.fcf <- ggplot(notath_scaled, 
       aes(x = body_index_notath, y = psite_count, colour = ci)) + geom_point() + xlab("FCF")  + 
  ylab("psite count")  + theme_bw() + stat_smooth(method=lm)

PV.psc.fcf <- ggplot(pimvig_scaled, 
       aes(x = body_index_pimvig, y = psite_count, colour = ci)) + geom_point() + xlab("FCF")  + 
  ylab("psite count")  + theme_bw() + stat_smooth(method=lm)

ggarrange(IP.psc.fcf, NA.psc.fcf, PV.psc.fcf, common.legend = TRUE, legend = "bottom", labels = c("A", "B", "C"))
# more variability than the others

```

### Plot life stage datausing scaled dataframe (with above-mentioned exclusions)
```{r life stage plots scaled, echo=TRUE}
# plot condition factor vs binary code for life stage (aka "stage-specific parasite stuff")
NA.bc.a <- ggplot(notath_scaled,
       aes(x = adult, y = body_index_ath, color = sex)) + xlab("Adult")  + geom_point() +
  ylab("Fulton's Condition Factor")  + theme_bw() + stat_smooth(method=lm)

NA.bc.l <- ggplot(notath_scaled,
       aes(x = larval, y = body_index_ath, color = sex)) + xlab("Larval")  + geom_point() +
  ylab("Fulton's Condition Factor")  + theme_bw() + stat_smooth(method=lm)

# plot condition factor vs binary code for life stage (aka "stage-specific parasite stuff")
IP.bc.a <- ggplot(ictpun_scaled,
       aes(x = adult, y = body_index_pun, color = sex)) + xlab("Adult")  + geom_point() + 
  ylab("Fulton's Condition Factor")  + theme_bw() + stat_smooth(method=lm)

IP.bc.l <- ggplot(ictpun_scaled, 
       aes(x = larval, y = body_index_pun, color = sex)) + xlab("Larval")  + geom_point() +
  ylab("Fulton's Condition Factor")  + theme_bw() + stat_smooth(method=lm)

# plot condition factor vs binary code for life stage (aka "stage-specific parasite stuff")
PV.bc.a <- ggplot(pimvig_scaled, 
       aes(x = adult, y = body_index_vig, color = sex)) + xlab("Adult")  + geom_point() +
  ylab("Fulton's Condition Factor")  + theme_bw() + stat_smooth(method=lm)

PV.bc.l <- ggplot(pimvig_scaled,
       aes(x = larval, y = body_index_vig, color = sex)) + xlab("Larval")  + geom_point() +
  ylab("Fulton's Condition Factor")  + theme_bw() + stat_smooth(method=lm)

ggarrange(IP.bc.a, IP.bc.l, NA.bc.a, NA.bc.l, PV.bc.a, PV.bc.a, common.legend = TRUE, legend = "bottom", labels = c("A", "B", "C", "D", "E", "F"))
# more variability than the others

## double check sex, weights
## should we exclude these outliers

```

## Models
We want to run a general linear model to test the research questions:
1. Is body condition associated with increased total parasite burden?
  - is there a linear association between increasing body condition and increasing parasite burden?
  - does FCF increase as parasite burden increases?
2. Is body condition associated with infection with adult or larval stages of parasites?
  - is there a linear association between increasing body condition and increasing abundance of larval or adult parasites in the host?
  - does FCF increase as adult or larval parasite presence increases?
  - variables: 
    - predictor = parasite_count, life_stage --> discrete
    - response = FCF --> continuous
  - discrete predictor = ANOVA
  - multiple/ continuous and discrete predictors = ANCOVA? 

#### Test Model Formula for question 1
$$
K_i\sim\alpha+\beta_1count+\epsilon_i
$$

#### Test Model Formula for question 2
$$
K_i\sim\alpha+\beta_1stage+\beta_2count+\epsilon_i
$$
### Run linear model to test above questions
```{r model}

# Question 1 (lm)
summary(FCF_ath_lm <- lm(body_index_ath ~ psite_count, data = Nath.life.stage))
plot(FCF_ath_lm)

summary(FCF_ath_lm <- lm(body_index_pun ~ psite_count, data = Ipun.life.stage))
plot(FCF_ath_lm)

summary(FCF_ath_lm <- lm(body_index_vig ~ psite_count, data = Pvig.life.stage))
plot(FCF_ath_lm)

# Question 1 (ANOVA)
summary(FCF_ath_anova <- aov(body_index_ath ~ psite_count, data = Nath.life.stage))
plot(FCF_ath_anova)

summary(FCF_ath_anova <- aov(body_index_pun ~ psite_count, data = Ipun.life.stage))
plot(FCF_ath_anova)

summary(FCF_ath_anova <- aov(body_index_vig ~ psite_count, data = Pvig.life.stage))
plot(FCF_ath_anova)


# Question 2 (lm)
summary(FCF_ath_al_lm <- lm(body_index_ath ~ larval * adult * psite_count, data = Nath.life.stage))
plot(FCF_ath_al_lm)

summary(FCF_ath_al_lm <- lm(body_index_pun ~ larval * adult * psite_count, data = Ipun.life.stage))
plot(FCF_ath_al_lm)

summary(FCF_ath_al_lm <- lm(body_index_vig ~ larval * adult * psite_count, data = Pvig.life.stage))
plot(FCF_ath_al_lm)

# Question 2 (ANOVA)
summary(FCF_ath_al_anova <- aov(body_index_ath ~ larval * adult * psite_count, data = Nath.life.stage))
plot(FCF_ath_al_anova)

summary(FCF_ath_al_anova <- aov(body_index_pun ~ larval * adult * psite_count, data = Ipun.life.stage))
plot(FCF_ath_al_anova)

summary(FCF_ath_al_anova <- aov(body_index_vig ~ larval * adult * psite_count, data = Pvig.life.stage))
plot(FCF_ath_al_anova)

```

# Full dataset analyses

#### Import dataset 
- make sure you name all these new datasets and frames something different from what you named them above (that way we can still run all the analyses in total without losing any of the previous ones)
```{r full dataset import, echo=TRUE}
full.data <- read.csv("data/processed/Full_dataset_with_psite_life_history_info_2024.07.25.csv") %>% 
  janitor::clean_names()
```

### Split everything into adult or larval based on the designations from initial analyses. Remove all non-concern cases of parasites (leech, nematomorph, copepod)
```{r adult vs larval splits}
full.data.stages <- full.data %>%
  mutate(life_stage = ifelse(psite_spp_x %in% c("META.HS",
                      "META.UNK", 
                      "MYX.GO",
                      "MYX.GEOM",
                      "MYXO.BC",
                      "MYX.THEL",
                      "MYXO.SBAD",
                      "MYXO.UNK",
                      "MYXO.UP",
                      "TREM.MET",
                      "TREM.SSS",
                      "TREM.BUC",
                      "TREM.METAF",
                      "TREM.METAGS",
                      "TREM.MS",
                      "TREM.NEA", # P. vigilax ends
                      "NEM.CYST",
                      "TREM.META.GG",
                      "TREM.META.UNK",
                      "TREM.LG",
                      "NEM.CYST",
                      "MYX.TAIL",
                      "MYX.GEOM", # I. punctatus ends
                      "NEM.CYST",
                      "TREM.LARV",
                      "TREM.META",
                      "MYX.SP", 
                      "MYX.ROUND",
                      "TREM.CYST", # N. atherinoides ends
                      "NEM.DAFT",
                      "TREM.CM",
                      "TREM.META.HET",
                      "TREM.META.SP",
                      "TREM.BC",
                      "TREM.META.GO",
                      "TREM.META.ES",
                      "META.UNK",
                      "TREM.META",
                      "TREM.META.UNK"), # H. nuchalis ends
                            "Larval", "Adult")) %>%
  janitor::clean_names()
    ```

- this is your code from before --> you could use this as a template for doing these analyses again with the new dataset
- use the distinctions we mapped out in the analyses above for what is adult or larval, since I think we have made some changes since this last one 
```{r}
Ipun.life.stage <- I.punctatus.data %>%
  mutate(LifeStage = ifelse(psite_spp %in% c("NEM.CYST",
                                             "TREM.LG",
                                             "TREM.META", # is this the same as META.GG - in data as TREM.META.GG
                                             "META.GG", # is this the same as TREM.META - in data as TREM.META.GG
                                             "META.UNK", # is this TREM.META.UNK in the data?
                                             "MYX.TAIL",
                                             "MYX.GEOM",
                                             "CYST.MX",
                                             "CYST.BD",
                                             "CYST.BLACK",
                                             "TREM.META.UNK"),
                            "Larval", "Adult")) # switching because there are more adults than larvae - less code to write
    # removing the species I'm not using
Ipun.life.stage <- subset(Ipun.life.stage, !(psite_spp %in% c("NMORPH",
                                                              "TREM.LARV",
                                                              "TREM.MID",
                                                              "CYST.UNKN", # correct code? This is what is in the slides
                                                              "WORM.UNK.B",
                                                              "TREM.2", # exclude because it's a once-off?
                                                              "LEECH.KUT"
                                                              )))

Nath.life.stage <- N.atherinoides.data %>%
  mutate(LifeStage = ifelse(psite_spp %in% c("NEM.CYST", # go through and check to see if dataset has any other not in the slides
                                             "TREM.LARV",
                                             "TREM.META",
                                             "MYX.SP", # MYX.SP.GILL?
                                             "MYX.DIAT",
                                             "MYX.ROUND",
                                             "CYST.UNA", # include?
                                             "CYST.UNK", # include?
                                             "TREM.CYST"), 
                            
                            "Larval", "Adult")) # switching because there are more adults than larvae - less code to write
# removing the species I'm not using
Nath.life.stage <- subset(Nath.life.stage, !(psite_spp %in% c("CYST.NEH",
                                                              "COPE.A",
                                                              "CYST.SAC"
)))

Pvig.life.stage <- P.vigilax.data %>%
  mutate(LifeStage = ifelse(psite_spp %in% c("ACANTH.BIGB", 
                                               "ACANTH.BKR",
                                               "NEM.CONA",
                                               "NEM.DICH",
                                               "NEM.LARV",
                                               "NEM.TRBK",
                                               "NEM.UNK",
                                               "CEST.GODZ",
                                               "CEST.NTG",
                                               "CEST.L",
                                               "CEST.VIT",
                                               "CEST.UNK",
                                               "MONO.GYRO",
                                               "MONO.DACT",
                                               "MONO.LG",
                                               "TREM.RNDA",
                                               "TREM.UNK"),
                            "Adult", "Larval"))
    # removing the species I'm not using
Pvig.life.stage <- subset(Pvig.life.stage, !(psite_spp %in% c("COPE.CL",
                                                      "COPE.GRAB",
                                                      "CYST.EMPTY",
                                                      "CYST.NEH",
                                                      "CYST.NP",
                                                      "WORM.A")))
```


### Remove from the dataframe the parasites that can only occur as either an adult or a larvae in the host. 
- acanths
- monogenes
- myxozoans 
Make sure you rename the new dataframe as something different from what you named it in the previous one (so that we can keep both versions to reference back to)
```{r only a/l both in host}
full.data.revised <- full.data.stages %>%
                              filter(!str_detect(psite_spp_x, "^(ACAN|MONO|COPE|LEECH|NMORPH|MYX)")) %>%
  janitor::clean_names()

# did not change psite_spp_y because it deletes some of the data when the y column equals NA: filter(!str_detect(psite_spp_y, "^(ACAN|MONO|COPE|LEECH|NMORPH|MYX)")) %>% 
# appears to be fine without it - checked using any() and checked the lengths of both columns
```

### Add back a column for Fulton's Condition Factor. 
```{r FCF into new dataframe}
full.data.revised <- full.data.revised %>%
  mutate(FCF = ((full.data.revised$weight_mg/((full.data.revised$standard_length_mm)^3)))*(100)) %>%
  janitor::clean_names() %>%
  filter(sex != "MA" & sex != "")
```


### Pivot and sum all the adult and larval stages as Chelsea described.
- you might need to play around with this section a bit
```{r add sums of parasites for each stage}
df <- df %>%
  group_by(fish_ID, psite_group, life_stage) %>%
  sum(???)
```

H. nuchalis parasite species to include w/ life stage:
A CEST.BB
A NEM.BUD
L NEM.DAFT
A NEM.UNK
L TREM.CM
A TREM.ASS
A TREM.NS
L TREM.META.HET
L TREM.META.SP
L TREM.BC
L TREM.META.GO
A TREM.DIPLO
L TREM.META.ES
A TREM.UNK
A CEST.UNK
L META.UNK
L TREM.META
L TREM.META.UNK


Species to remove:
MYX.AK
MYX.FI
MYX.OT
MYX.TIN
MYX.GL
MYX.EYE
MYX.GC
MONO.DACT
MONO.GDAC
ACAN.AD - one-off (occurs 4 times in all species)
COPE.POOD - one-off (occurs 3 times in all species)



```{r dont forget to have fun!}

face <- data.frame(x = 0, y = 0, r = 10)
left_eye <- data.frame(x = -3, y = 3, r = 1.5)
right_eye <- data.frame(x = 3, y = 3, r = 1.5)
mouth <- data.frame(
  x = c(-5, -4, -3, -2, -1, 0, 1, 2, 3, 4, 5),
  y = c(0, -1, -2, -2.5, -3, -3, -3, -2.5, -2, -1, 0)
)
ggplot() +
  geom_point(data = face, aes(x = x, y = y), size = 100, shape = 21, fill = "yellow", color = "black") +
  geom_point(data = left_eye, aes(x = x, y = y), size = 20, shape = 21, fill = "white", color = "black") +
  geom_point(data = left_eye, aes(x = x, y = y), size = 10, shape = 21, fill = "black", color = "black") +
  geom_point(data = right_eye, aes(x = x, y = y), size = 20, shape = 21, fill = "white", color = "black") +
  geom_point(data = right_eye, aes(x = x, y = y), size = 10, shape = 21, fill = "black", color = "black") +
  geom_line(data = mouth, aes(x = x, y = y), size = 2, color = "black") +
  coord_fixed() +
  theme_void() +
  theme(panel.background = element_rect(fill = "skyblue", color = "skyblue"))

```

